Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-26 paper/ijcai16.tex

We adopt the {\em indifference-augmented} Bradley-Torr user response model
introduced in~\cite{guo2010real}. When queried about a pair of configurations,
the user replies with the true utility of the two items. According to the
classical Bradley-Torr model, which does not attempt to model {\em indifference},
the probability that a user $\vw$ will prefer configuration $\vx^i$ over
the alternative $\vx^j$, barring indifference, amounts to:
%
$$ (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
%
In other words, the probability grows with the difference between the utilities
of the two items. Support for indifference has been introduced
in~\cite{guo2010real} by postulating that the probability of the user being
indifferent about the two configurations $\vx^i, \vx^j$ depends on how close
the two utilities are, as in:
%
$$ \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
%
The parameters $\alpha$ and $\beta$ were set as in~\cite{guo2010real}. Please
see the latter for more details
\stefano{The above does not satisfy me at all.}

In all experiments we use an internal 5-fold cross-validation procedure to
update the hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5
iterations. The hyperparameters are chosen as to minimize the ranking loss over
the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.
\stefano{explain that we can't turn off $\gamma$}

\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
methods, which generate a single candidate option per iteration, {\sc
SetMargin} constructs a whole {\em set} of options. We thus can not leverage
standard query selection strategies \stefano{which ones?}. 
\paolo{I would remove the last sentence}

\paragraph{Synthetic Dataset.} Following the experimental protocol
in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first experiment
we evaluate the behavior of the proposed method in a challenging artificial
setting. We developed seven synthetic datasets: each dataset involves
$m=2,\ldots,7$ attributes, where each attribute takes one of $m$ possible
values. More explicitly, for $m=2$ the synthetic dataset is
$\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to $\calX_\text{feasible} =
[3] \times [3] \times [3]$, and so on. The cardinality of
$\calX_\text{feasible}$ is $m^m$, and grows (super) exponentially with $m$. For
$m=3$, the dataset is comparable to the synthetic one used
in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
of size of 2, 2 and 5, respectively, for a total of 20 feasible
configurations), while for larger $m$ it grows much larger than those typically
used in the Bayesian preference elicitation literature, and as such represents
a good test bed for comparing the scalability of the various methods.
The feasible configuration space was encoded in {\sc SetMargin} through
appropriate MILP constraints, while the other methods required all datasets to
be explicitly grounded.

@@

In Figure~\ref{fig:} we plot, for each setting, the behavior of the utility
loss:
%
$$ \left(\max_{\vx\in\calX_\text{feasible}} u(\vx)) - u(\vx^{i}) \right) $$
%
varying the number of collected user responses. The thick line is the median
utility loss taken over all 20 users, while the shaded area represents the
variance.

@@

@@

A PC configurations is defined by seven attributes: computer type (laptop,
desktop, or tower), manufacturer, CPU model, monitor size, RAM amount, storage
size, and price. See Table~\ref{tab:pcdataset} for an overivew of the
attributes. The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by sum of the price of its components plus an offset due to
branding. Compared to the PC dataset in~\cite{guo2010real}, we combined the CPU
type and CPU frequency attributes into a single CPU attribute (e.g. ``PowerPC
G4 @733 MHz''), so to avoid multiplicative terms in the price function.
\stefano{too detailed/handwavy?}

Interactions between attributes are expressed as Horn clauses. An example
costraint might look like this: ``if the manufacturer is Apple, then the
CPU must be either a PowerPC G3 or a PowerPC G4'', which can be encoded
into MILP form as:
%
$$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $$
%
Note that mutual exclusivity between all binary variables corresponding to an
attribute ensures that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1.
The dataset includes constraints between the following attributes: manufacturer
$\to$ type, manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
size, type $\to$ monitor size. In total we encoded 16 Horn constraints, each
with one head and multiple options in the body. We do not report the full
list due to space limitations.

>>>>

We adopt the {\em indifference-augmented} Bradley-Terry user response
model introduced in~\cite{guo2010real}. According to the classical
(without indifference) Bradley-Terry model~\cite{BraTer52}, when queried about a pair
of configurations, the probability that a user will prefer
configuration $\vx^i$ over the alternative $\vx^j$ amounts to:

%
$$ (1 + \exp(-\lambda_1 \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
%
where $\vw$ is the weight vector of the true underlying user utility.
In other words, the probability grows with the difference between the
utilities of the two items. Support for indifference has been
introduced in~\cite{guo2010real} by postulating that the probability
of the user being indifferent about the two configurations depends on
how close their utilities are, as in:
%
$$ \exp(\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|) $$
%
The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
simulations, as in~\cite{guo2010real}.

In all experiments we use an internal 5-fold cross-validation procedure to
update the hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5
iterations. The hyperparameters are chosen as to minimize the ranking loss over
the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
% <<<<<<< HEAD
% while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.
% \stefano{explain that we can't turn off $\gamma$}
% 
% \paragraph{Cognitive effort.} Contrarily to most other preference elicitation
% methods, which generate a single candidate option per iteration, {\sc
% SetMargin} constructs a whole {\em set} of options. We thus can not leverage
% standard query selection strategies \stefano{which ones?}. 
% \paolo{I would remove the last sentence}
% 
% The simplest solution is to query the user about all $k \choose 2$ pairs of
% configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
% number of queries does not grow large in practice.  An appealing alternative
% involves using a modified sorting algorithm, such as merge-sort, to order the
% set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
% approach is very susceptible to noise: a single inconsistent user response may
% affect all successive comparisons, leading to a cascading effect. While more
% robust strategies may be conceived, empirically we found that the simpler
% quadratic scheme tends to work reasonably well.
% 
% \andrea{rewrite as footnote}
% =======
while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$\footnote{Note
that $\alpha$ can not be taken to be less than $1$, as in this case, the
objective can be increased arbitrarily while keeping the right-hand side of
Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}.

\paragraph{Synthetic Dataset.} Following the experimental protocol
in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first
experiment we evaluate the behavior of the proposed method in an
artificial setting with increasingly complex problems. We developed
synthetic datasets with $r$ attributes, for increasing values of $r$.
Each attribute takes one of $r$ possible values, so that the one-hot
encoding of attributes results in $m=r^2$ features. In terms of space
of configurations, for $r=3$ the synthetic dataset corresponds to 
$\calX_\text{feasible} = [3] \times [3] \times [3]$, for $r=4$ to
$\calX_\text{feasible} = [4] \times [4] \times [4] \times [4]$, and so on. The
cardinality of $\calX_\text{feasible}$ is $r^r$, and grows (super)
exponentially with $r$.
% We developed
% seven synthetic datasets: each dataset involves $m=2,\ldots,7$
% attributes, where each attribute takes one of $m$ possible
% values. More explicitly, for $m=2$ the synthetic dataset is
% $\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to
% $\calX_\text{feasible} = [3] \times [3] \times [3]$, and so on. The
% cardinality of $\calX_\text{feasible}$ is $m^m$, and grows (super)
% exponentially with $m$. 
For $r=3$, the dataset is comparable to the synthetic one used
in~\cite{guo2010real} and~\cite{viappiani2010optimal}, which have
three attributes with 2, 2 and 5 values respectively, for a total of
20 feasible configurations.  For larger $r$ the size of the space
grows much larger than the one typically used in the Bayesian
preference elicitation literature, and as such represents a good test
bed for comparing the scalability of the various methods. The feasible
configuration space was encoded in {\sc SetMargin} through appropriate
MILP constraints, while the other methods require all datasets to be
explicitly grounded.

@@

In Figure~\ref{fig:} we plot, for each setting, the behavior of the utility
loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
varying the number of collected user responses. Here $u(\cdot)$ is the
true unknown user utility, and $\vx^*$ is the solution recommended to
the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).

The thick line is the median
utility loss taken over all 20 users, while the shaded area represents the
standard deviation.

@@

\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
methods, which generate a single candidate option per iteration, {\sc
SetMargin} constructs a whole {\em set} of options. We thus can not leverage
standard query selection strategies \stefano{which ones?}.

@@

A PC configurations is defined by seven attributes: computer type (laptop,
desktop, or tower), manufacturer, CPU model, monitor size, RAM amount, storage
size, and price. See Table~\ref{tab:pcdataset} for an overivew of the
attributes. The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by sum of the price of its components plus an offset due to
branding.

Interactions between attributes are expressed as Horn clauses. An example
constraint might look like this: ``if the manufacturer is Apple, then the
CPU must be either a PowerPC G3 or a PowerPC G4''\footnote{Despite the components
included in the dataset, which are a bit outdated, the dataset itself is pretty
realistic.}, which can be encoded
into MILP form as:
%
$$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $$
%
Note that mutual exclusivity between all binary variables of the
one-hot encoding of an attribute ensures that only one of
$x_\text{G3}$ and $x_\text{G4}$ can be 1.  The dataset includes
constraints between the following attributes: manufacturer $\to$ type,
manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
size, type $\to$ monitor size. In total we encoded 16 Horn
constraints, each with one head and multiple options in the body. We
do not report the full list due to space limitations.
